---
title: "Lake Erie Microcystin Risk Assessment -- Model Averaging"
author: "Song Qian"
date: "8/15/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initial setup
```{r}
source("FrontMatter.R")
packages(loo)
load(paste(saveDIR, "eriedata2019.RData", sep="/"))
load(paste(saveDIR, "eriedata2021.RData", sep="/"))
source("sequentialANN.R")## year-end updating 
source("sequentialSN.R") 
source("sequentialSN_BMA.R") ## functions for input data
## R functions for 
## sub-setting data
## prior functions
## Stan models and input functions
## compiling Stan models -- need to recompile from time to time
## within-year updating without season
## stan_mod_no_cens <- stan_model(model_code = stan_model1)
## stan_mod_cens <- stan_model(model_code = stan_model1_cens)
## stan_mod_allcens <- stan_model(model_code = stan_model1_5)
## stan_mod_add_no_cens <- stan_model(model_code = stan_model3)
## stan_mod_add_cens <- stan_model(model_code = stan_model3_cens)
## stan_mod_add_allcens <- stan_model(model_code = stan_model3_5)
## within-year undating nested
## stan_mod_nstd_no_cens <- stan_model(model_code = stan_model4)
## stan_mod_nstd_cens <- stan_model(model_code = stan_model4_cens)
## stan_mod_nstd_allcens <- stan_model(model_code = stan_model4_5)
## year-end updating
## stan_mod_ann_add <- stan_model(model_code = stan_model5_add) 
## stan_mod_ann_nstd <- stan_model(model_code = stan_model5_nstd) 

##save(stan_mod_no_cens, stan_mod_cens, stan_mod_allcens, ##stan_mod_add_no_cens,stan_mod_add_cens,stan_mod_add_allcens,stan_mod_nstd_no_cens,stan_mod_nstd_cens,stan_mod_##nstd_allcens,stan_mod_ann_add, stan_mod_ann_nstd, file=paste(saveDIR, "compiledStan.RData", sep="/"))
## annual updating model with season
load(paste(saveDIR, "compiledStan.RData", sep="/"))
niters <- 50000
nkeep <- 2500
nthin <- ceiling((niters/2)*nchains/nkeep)
```

## Annual Data
Simulating biweekly sampling schedule

```{r}
## Weekly updating function
seq_within <- function(data=tempdata, 
                       priorAdd_R=prior.rv, priorAdd_Sn=priorSn.rv[[1]], 
                       priorNstd_R=priorYr.rvnstd, priorNstd_Sn=priorSn.rvnstd, 
                       prior2W_R=priorYr.rv2W, prior2W_Sn=priorSn.rv2W, 
                       priorNSN=prior.rv){ 
  ## within-year one-step updating function
  #### 1. running additive updating

  input.to.stan <- stan.in_seq_add(infile=data, x="Chla", y="part_microcystin",
                                   n.chains=nchains, priorYr_rv=priorAdd_R,
                                   priorSn_rv=priorAdd_Sn)
  if (input.to.stan$cens[1] == 0) {
    stan_mod <- stan_mod_add_allcens
  } else if(input.to.stan$cens[2] == 0){
    stan_mod <- stan_mod_add_no_cens
  } else {
    stan_mod <- stan_mod_add_cens
  }
  ## additive model
  print("Now fitting the additive model")
  fit1 <- sampling(stan_mod,
         			     data = input.to.stan$data,
	                 init = input.to.stan$inits,
                   pars = input.to.stan$para,
                   iter = niters, thin=nthin, 
                   chains=input.to.stan$n.chains, 
	                 control = list(adapt_delta = 0.9, max_treedepth=15))
  fitcoef1 <-
    rvsims(as.matrix(as.data.frame(rstan::extract(fit1, permute=T))))
  prr_yr1 <- list(B0=fitcoef1["beta0"], De=fitcoef1["delta"], Ph=fitcoef1["phi"])
    ## feeding to input function "priorYr_rv"
  prr_sn1 <- list(d0=fitcoef1["d0"], dp=fitcoef1["dp"], dd=fitcoef1["dd"])
  log_lik_1 <- extract_log_lik(fit1, merge_chains = FALSE)
  r_eff_1 <- relative_eff(exp(log_lik_1))
  loo_1 <- loo(log_lik_1, r_eff = r_eff_1, cores = 2)
  
  #### 2. nested (by month)
  print("Now fitting the nested (by month) model")
  input.to.stan <- stan.in_seq_nst(infile=data, x="Chla", y="part_microcystin",
                                   n.chains=nchains, priorYr_rv=priorNstd_R,
                                   priorSn_rv=priorNstd_Sn)
  if (input.to.stan$cens[1] == 0) {
    stan_mod <- stan_mod_nstd_allcens
  } else if(input.to.stan$cens[2] == 0){
    stan_mod <- stan_mod_nstd_no_cens
  } else {
    stan_mod <- stan_mod_nstd_cens
  }
  fit2 <- sampling(stan_mod,
         			     data = input.to.stan$data,
	                 init = input.to.stan$inits,
                   pars = input.to.stan$para,
                   iter = niters, thin=nthin, 
                   chains=input.to.stan$n.chains, 
	                 control = list(adapt_delta = 0.9, max_treedepth=15))
  fitcoef2 <-
    rvsims(as.matrix(as.data.frame(rstan::extract(fit2, permute=T))))
  prr_yr2 <- list(B0=fitcoef2["beta0"], De=fitcoef2["delta"], Ph=fitcoef2["phi"])
    ## feeding to input function "priorYr_rv"
  prr_sn2 <- list(sigmaS0=fitcoef2["sigma_d0"], sigmaSd=fitcoef2["sigma_dd"], sigmaSp=fitcoef2["sigma_dp"])

  log_lik_2 <- extract_log_lik(fit2, merge_chains = FALSE)
  r_eff_2 <- relative_eff(exp(log_lik_2))
  loo_2 <- loo(log_lik_2, r_eff = r_eff_2, cores = 2)
  
  #### 3. nested (biweekly)
  print("Now fitting the nested (2w) model")
  input.to.stan <- stan.in_seq_nst(infile=data, x="Chla", y="part_microcystin",
                                   n.chains=nchains, priorYr_rv=prior2W_R,
                                   priorSn_rv=prior2W_Sn)
  if (input.to.stan$cens[1] == 0) {
    stan_mod <- stan_mod_nstd_allcens
  } else if(input.to.stan$cens[2] == 0){
    stan_mod <- stan_mod_nstd_no_cens
  } else {
    stan_mod <- stan_mod_nstd_cens
  }
  fit3 <- sampling(stan_mod,
         			     data = input.to.stan$data,
	                 init = input.to.stan$inits,
                   pars = input.to.stan$para,
                   iter = niters, thin=nthin, 
                   chains=input.to.stan$n.chains, 
	                 control = list(adapt_delta = 0.9, max_treedepth=15))
  fitcoef3 <-
    rvsims(as.matrix(as.data.frame(rstan::extract(fit3, permute=T))))
  prr_yr3 <- list(B0=fitcoef3["beta0"], De=fitcoef3["delta"], Ph=fitcoef3["phi"],
                  sigmaR0=fitcoef3["sigma_d0"], sigmaRd=fitcoef3["sigma_dd"], 
                  sigmaRp=fitcoef3["sigma_dp"])
    ## feeding to input function "priorYr_rv"
  prr_sn3 <- list(sigmaS0=fitcoef3["sigma_d0"], sigmaSd=fitcoef3["sigma_dd"], sigmaSp=fitcoef3["sigma_dp"])
  log_lik_3 <- extract_log_lik(fit3, merge_chains = FALSE)
  r_eff_3 <- relative_eff(exp(log_lik_3))
  loo_3 <- loo(log_lik_3, r_eff = r_eff_3, cores = 2)
  
  #### 4. no season
  print("Now fitting the no-season model")
  input.to.stan <- stan.in_seq_nsn(infile=data, x="Chla", y="part_microcystin",
                                   n.chains=nchains, priorYr_rv=priorNSN)
  if (input.to.stan$cens[1] == 0) {
    stan_mod <- stan_mod_allcens
  } else if(input.to.stan$cens[2] == 0){
    stan_mod <- stan_mod_no_cens
  } else {
    stan_mod <- stan_mod_cens
  }
  fit4 <- sampling(stan_mod,
         			     data = input.to.stan$data,
	                 init = input.to.stan$inits,
                   pars = input.to.stan$para,
                   iter = niters, thin=nthin, 
                   chains=input.to.stan$n.chains, 
                   control = list(adapt_delta = 0.9, max_treedepth=15))
  fitcoef4 <-
    rvsims(as.matrix(as.data.frame(rstan::extract(fit4, permute=T))))
  prr_yr4 <- list(B0=fitcoef4["mu0"], De=fitcoef4["muD"], Ph=fitcoef4["muP"], 
                  sigma0=fitcoef4["sigma0"], sigmad=fitcoef4["sigmad"], 
                  sigmap=fitcoef4["sigmap"])
    ## feeding to input function "priorYr_rv"
  log_lik_4 <- extract_log_lik(fit4, merge_chains = FALSE)
  r_eff_4 <- relative_eff(exp(log_lik_4))
  loo_4 <- loo(log_lik_4, r_eff = r_eff_4, cores = 2)
  lpd_point <- cbind(
    loo_1$pointwise[,"elpd_loo"], ## additive
    loo_2$pointwise[,"elpd_loo"], ## nested monthly
    loo_3$pointwise[,"elpd_loo"], ## nested biweekly
    loo_4$pointwise[,"elpd_loo"]  ## no season
  )
  return(list(fits=list(fit1, fit2, fit3, fit4), 
              wts = list(pbma_wts=pseudobma_weights(lpd_point, BB=FALSE), 
                         pbma_BB_wts=pseudobma_weights(lpd_point),  # default is BB=TRUE
                         stacking_wts=stacking_weights(lpd_point)), 
              priors=list(add_yr=prr_yr1, add_sn=prr_sn1, 
                          nstd_yr=prr_yr2, nstd_sn=prr_sn2, 
                          n2w_yr=prr_yr3, n2w_sn=prr_sn3, 
                          nsn_yr=prr_yr4)))
}

```

## Sequential updating within a year

Each year's data file is grouped by sampling period and month. Posteriors of the previous period are used to derive priors for the next period.

### Function 
```{R}
## using the function `seq_within()`

seq_within_allY <- function(data=tmpdata, add_yr0=priorYr.add, add_sn0=priorSn.add,
                            nstd_yr0=priorYr.nstd, nstd_sn0=priorSn.nstd, 
                            n2w_yr0=priorYr.2W, n2w_sn0 = priorSn.2W, 
                            nsn_yr0=priors.nsn){
wks <- as.numeric(ordered(data$BiWeek))
wksL <- as.numeric(levels(ordered(data$BiWeek)))
add_yr=add_yr0
add_sn=add_sn0
nstd_yr=nstd_yr0
nstd_sn=nstd_sn0=priorSn.nstd
n2w_yr=n2w_yr0=priorYr.2W
n2w_sn=n2w_sn0 
nsn_yr=nsn_yr0
seq_out <- list()
for (i in 1:max(wks)){
  print(paste("Iteration ", i, "of ", max(wks)))
  tempdata <- data[data$BiWeek==wksL[i], ] 
  mns <- tempdata$Mon_ord[i]
  if (i == 1){
    add_sn <- add_sn0[[mns]]
  } else if (mns != tempdata$Mon_ord[i-1])
    add_sn <- add_sn0[[mns]]
## data for each within-year updating
## one updating step: (1) processing priors and input data for 4 models
## one-time within-year updating
seq_out[[i]]<-seq_within(data=tempdata, priorAdd_R=add_yr, priorAdd_Sn=add_sn, 
                         priorNstd_R=nstd_yr, priorNstd_Sn=nstd_sn, 
                         prior2W_R=n2w_yr, prior2W_Sn=n2w_sn, priorNSN=nsn_yr)
add_yr$B0 <- seq_out[[i]]$priors$add_yr$B0
add_yr$De <- seq_out[[i]]$priors$add_yr$De
add_yr$Ph <- seq_out[[i]]$priors$add_yr$Ph
add_sn <- seq_out[[i]]$priors$add_sn
nstd_yr$B0 <- seq_out[[i]]$priors$nstd_yr$B0
nstd_yr$De <- seq_out[[i]]$priors$nstd_yr$De
nstd_yr$Ph <- seq_out[[i]]$priors$nstd_yr$Ph
nstd_sn <- seq_out[[i]]$priors$nstd_sn
n2w_yr$B0 <- seq_out[[i]]$priors$n2w_yr$B0
n2w_yr$De <- seq_out[[i]]$priors$n2w_yr$De
n2w_yr$Ph <- seq_out[[i]]$priors$n2w_yr$Ph
n2w_sn <- seq_out[[i]]$priors$n2w_sn
nsn_yr <- seq_out[[i]]$priors$nsn_yr
}
return(list(fitted=seq_out, priors=list(addYr=add_yr, addSn=add_sn, nstdYr=nstd_yr, 
                                        nstdSn = nstd_sn, n2wYr=n2w_yr, n2wSn=n2w_sn)))
}

```

## Sequential Updating
For each year, we break the data into updating groups and update the model once each sampling event concluded.

Starting from the end of year 2018
```{r}
## extracting priors for 2019
## Function to derive prior distribution parameters from a Stan BHM out file.
load(paste(saveDIR, "MC_ChlaCens.RData", sep="/")) ## starting from 2019 with seasonal terms
load(paste(saveDIR, "MC_ChlaCens_2019.RData", sep="/")) ## without seasonal terms
print(fit2keep_full_2019$fit)  ## without season
print(fit2keep_full_add$fit)   ## additive
print(fit2keep_full_nstd$fit)  ## nested month
print(fit2keep_full_2W$fit)    ## nested biweekly

stanfit_full_2019 <- rstan::extract(fit2keep_full_2019$fit)
stanfit_full_add <- rstan::extract(fit2keep_full_add$fit)
stanfit_full_nstd <- rstan::extract(fit2keep_full_nstd$fit)
stanfit_full_2W <- rstan::extract(fit2keep_full_2W$fit)

## extracting priors
#### without seasons
prrs_nsn <- rvsims(as.matrix(as.data.frame(stanfit_full_2019))) 
priors.nsn <- list(B0=prrs_nsn["B0"], De=prrs_nsn["De"], Ph=prrs_nsn["Ph"], 
                   sigma0=prrs_nsn["sigma0"], sigmad=prrs_nsn["sigmad"], 
                   sigmap=prrs_nsn["sigmap"])
## prior parameters needed: m0;mD;mP;lmbd0;lmbdD;lmbdP;al0;alP;alD;bt0;btP;btD;
## all for N-IG prior

#### additive model
prrs_add <- rvsims(as.matrix(as.data.frame(stanfit_full_add))) 
## feeding to input function "priorYr_rv"
priorYr.add <- list(B0=prrs_add["B0"], De=prrs_add["De"], Ph=prrs_add["Ph"], 
                    sigmaR0=prrs_add["sigmaR0"], sigmaRd=prrs_add["sigmaRd"], 
                    sigmaRp=prrs_add["sigmaRp"], 
                    sigmaS0=prrs_add["sigmaS0"], sigmaSd=prrs_add["sigmaSd"], 
                    sigmaSp=prrs_add["sigmaSp"])
## sigmaRs needed for annual updating, but not for within a year

## feeding to input function "priorSn_rv", not needed for year-end updating
priorSn.add <- list(list(d0=prrs_add["delB0S.1"], dp=prrs_add["delPhS.1"], dd=prrs_add["delDS.1"]),
                    list(d0=prrs_add["delB0S.2"], dp=prrs_add["delPhS.2"], dd=prrs_add["delDS.2"]),
                    list(d0=prrs_add["delB0S.3"], dp=prrs_add["delPhS.3"], dd=prrs_add["delDS.3"]),
                    list(d0=prrs_add["delB0S.4"], dp=prrs_add["delPhS.4"], dd=prrs_add["delDS.4"]),
                    list(d0=prrs_add["delB0S.5"], dp=prrs_add["delPhS.5"], dd=prrs_add["delDS.5"]),
                    list(d0=prrs_add["delB0S.6"], dp=prrs_add["delPhS.6"], dd=prrs_add["delDS.6"]))
### !: ".1",".2"... for extracted stan object. "delBOS[1]" without extract

priorSn.add2019 <- priorSn.add ## save for model evaluation

#### nested version (monthly)
prrs_nstd <- rvsims(as.matrix(as.data.frame(stanfit_full_nstd))) 
priorYr.nstd <- list(B0=prrs_nstd["B0"], De=prrs_nstd["De"], Ph=prrs_nstd["Ph"], 
                    sigmaR0=prrs_nstd["sigmaR0"], sigmaRd=prrs_nstd["sigmaRd"], 
                    sigmaRp=prrs_nstd["sigmaRp"])
priorSn.nstd <- list(sigmaS0=prrs_nstd["sigmaS0"], sigmaSd=prrs_nstd["sigmaSd"], 
                     sigmaSp=prrs_nstd["sigmaSp"])
## nested -2 week interval
prrs_2W <- rvsims(as.matrix(as.data.frame(stanfit_full_2W))) 
priorYr.2W <- list(B0=prrs_2W["B0"], De=prrs_2W["De"], Ph=prrs_2W["Ph"], 
                   sigmaR0=prrs_2W["sigmaR0"], sigmaRd=prrs_2W["sigmaRd"], 
                   sigmaRp=prrs_2W["sigmaRp"])
priorSn.2W <- list(sigmaS0=prrs_2W["sigmaS0"], sigmaSd=prrs_2W["sigmaSd"], 
                   sigmaSp=prrs_2W["sigmaSp"])
```
## Annual updating

Now running sequential updating within a year (2019):
```{R}
## 2019
## processing data
load(paste(saveDIR, "eriedata2019.RData", sep="/"))

temp<-(!is.na(eriedata2019$part_microcystin))&(!is.na(eriedata2019$Chla))&(eriedata2019$Chla>0)& 
  (eriedata2019$part_microcystin>0)&(eriedata2019$Month>=5 & eriedata2019$Month<=10)
tmpdata <- eriedata2019[temp,]
tmpdata$Mon_ord <- as.numeric(ordered(tmpdata$Month, levels=5:10))
tmpdata$Yr_ord <- as.numeric(ordered(tmpdata$Year))
tmpdata$sn_nstd <- as.numeric(ordered(paste(tmpdata$Year, tmpdata$Month, sep=":")))
## for biweek season
tmpdata$sn_nstd2w <- as.numeric(ordered(paste(tmpdata$Year, tmpdata$BiWeek, sep=":")))

seq_2019 <- seq_within_allY ()
save("seq_2019", file=paste(saveDIR, "seq2019.RData", sep="/"))
## load(paste(saveDIR, "seq2019.RData", sep="/"))
```

At the end of the year, updating the BHM using data from the entire year.

```{R}
## running annual sequential updating
### extracting prior parameters first from the full model up to 2018
### these are the same priors used in the previous code chunk
### 1. Additive model
prrs_add <- rvsims(as.matrix(as.data.frame(stanfit_full_add))) 
## feeding to input function "priorYr_rv"
priorYr.add <- list(B0=prrs_add["B0"], De=prrs_add["De"], Ph=prrs_add["Ph"], 
                   sigmaR0=prrs_add["sigmaR0"], sigmaRd=prrs_add["sigmaRd"], 
                   sigmaRp=prrs_add["sigmaRp"], 
                   sigmaS0=prrs_add["sigmaS0"], sigmaSd=prrs_add["sigmaRd"], 
                   sigmaSp=prrs_add["sigmaSp"])
priorSn.add <- list(list(d0=prrs_add["delB0S.1"], dp=prrs_add["delPhS.1"], dd=prrs_add["delDS.1"]),
                    list(d0=prrs_add["delB0S.2"], dp=prrs_add["delPhS.2"], dd=prrs_add["delDS.2"]),
                    list(d0=prrs_add["delB0S.3"], dp=prrs_add["delPhS.3"], dd=prrs_add["delDS.3"]),
                    list(d0=prrs_add["delB0S.4"], dp=prrs_add["delPhS.4"], dd=prrs_add["delDS.4"]),
                    list(d0=prrs_add["delB0S.5"], dp=prrs_add["delPhS.5"], dd=prrs_add["delDS.5"]),
                    list(d0=prrs_add["delB0S.6"], dp=prrs_add["delPhS.6"], dd=prrs_add["delDS.6"]))

### 2. nested
prrs_nstd <- rvsims(as.matrix(as.data.frame(stanfit_full_nstd))) 
## feeding to input function "priorYr_rv"
priorYr.nstd <- list(B0=prrs_nstd["B0"], De=prrs_nstd["De"], Ph=prrs_nstd["Ph"], 
                     sigmaR0=prrs_nstd["sigmaR0"], sigmaRd=prrs_nstd["sigmaRd"], 
                     sigmaRp=prrs_nstd["sigmaRp"])
priorSn.nstd <- list(sigmaS0=prrs_add["sigmaS0"], sigmaSd=prrs_add["sigmaSd"], 
                     sigmaSp=prrs_add["sigmaSp"])

### 3. Biweekly
prrs_2w <- rvsims(as.matrix(as.data.frame(stanfit_full_2W))) 
## feeding to input function "priorYr_rv"
priorYr.2w <- list(B0=prrs_2w["B0"], De=prrs_2w["De"], Ph=prrs_2w["Ph"], 
                   sigmaR0=prrs_2w["sigmaR0"], sigmaRd=prrs_2w["sigmaRd"], 
                   sigmaRp=prrs_2w["sigmaRp"]) 
priorYr.2w <- list(sigmaS0=prrs_2w["sigmaS0"], sigmaSd=prrs_2w["sigmaSd"], 
                   sigmaSp=prrs_2w["sigmaSp"])
## sigmaRs needed for annual updating, but not for within a year

### 4. No season using `stan_mod_cens` (line 97)
prrs_nsn <- rvsims(as.matrix(as.data.frame(stanfit_full_2019))) 
priors.nsn <- list(B0=prrs_nsn["B0"], De=prrs_nsn["De"], Ph=prrs_nsn["Ph"], 
                   sigma0=prrs_nsn["sigma0"], sigmad=prrs_nsn["sigmad"], 
                   sigmap=prrs_nsn["sigmap"])
## prior parameters needed: m0;mD;mP;lmbd0;lmbdD;lmbdP;al0;alP;alD;bt0;btP;btD;
## all for N-IG prior

## 2019 year end
## processing data
tmpdata <- eriedata2019
temp<-(!is.na(tmpdata$part_microcystin))&(!is.na(tmpdata$Chla))&(tmpdata$Chla>0)& 
  (tmpdata$part_microcystin>0)&(tmpdata$Month>=5 & tmpdata$Month<=10)
tmpdata <- eriedata2019[temp,]
tmpdata$Mon_ord <- as.numeric(ordered(tmpdata$Month, levels=5:10))
tmpdata$Yr_ord <- as.numeric(ordered(tmpdata$Year))
tmpdata$sn_nstd <- as.numeric(ordered(paste(tmpdata$Year, tmpdata$Month, sep=":")))
## for biweek season
tmpdata$sn_nstd2w <- as.numeric(ordered(paste(tmpdata$Year, tmpdata$BiWeek, sep=":")))

## no seasons
input.to.stan <- stan.in_seq_nsn(infile=tmpdata, priorYr_rv = priors.nsn)
fitNSN2019 <- sampling(stan_mod_cens,
         			     data = input.to.stan$data,
	                 init = input.to.stan$inits,
                   pars = input.to.stan$para,
                   iter = niters, thin=nthin, 
                   chains=input.to.stan$n.chains) #, 
##                   control = list(adapt_delta = 0.9, max_treedepth=15))
## additive
input.to.stan <- stan.in_seqA_add(infile=tmpdata, x="Chla", y="part_microcystin",
                              n.chains=nchains, priorYr_rv=priorYr.add, 
                              priorSn_rv=priorSn.add)

fit2keep_2019_add <- sampling(stan_mod_ann_add, 
                              data = input.to.stan$data,
                              init=input.to.stan$inits,
                              pars = input.to.stan$para,
                              iter=niters, thin=nthin,
                              chains=input.to.stan$n.chains)

print(fit2keep_2019_add)
plot(fit2keep_2019_add)

## nested (month)
input.to.stan <- stan.in_seqA_nstd(infile=tmpdata, x="Chla", y="part_microcystin",
                              n.chains=nchains, priorYr_rv=priorYr.nstd, priorSn_rv=priorSn.nstd,biweek=F)

fit2keep_ann_2019N <- sampling(stan_mod_ann_nstd, 
                               data = input.to.stan$data,
                               init=input.to.stan$inits,  
                               pars = input.to.stan$para,
                               iter=niters, thin=nthin,
                               chains=input.to.stan$n.chains)#)

print(fit2keep_ann_2019N)
plot(fit2keep_ann_2019N)
## 2w
input.to.stan <- stan.in_seqA_nstd(infile=tmpdata, x="Chla", y="part_microcystin",
                                   n.chains=nchains, priorYr_rv=priorYr.nstd, 
                                   priorSn_rv=priorSn.nstd, biweek=T)

fit2keep_ann_2019W <- sampling(stan_mod_ann_nstd,  data = input.to.stan$data,
                               init=input.to.stan$inits,
                               pars = input.to.stan$para,
                               iter=niters, thin=nthin,
                               chains=input.to.stan$n.chains)#)

print(fit2keep_ann_2019W)
plot(fit2keep_ann_2019W)

save("fitNSN2019", "fit2keep_2019_add", "fit2keep_ann_2019N", "fit2keep_ann_2019W", 
     file=paste(saveDIR, "annual2019.RData", sep="/"))
## load(paste(saveDIR, "annual2019.RData", sep="/"))
```

Now running sequential updating within a year (2020):
```{R}
## 2020
## extracting priors
## updating priors for 2020:
### 1. Additive model
prrs_add <- rvsims(as.matrix(as.data.frame(rstan::extract(fit2keep_2019_add))))
## feeding to input function "priorYr_rv"
priorYr.add <- list(B0=prrs_add["B0"], De=prrs_add["De"], Ph=prrs_add["Ph"], 
                    sigmaR0=prrs_add["sigmaR0"], sigmaRd=prrs_add["sigmaRd"], 
                    sigmaRp=prrs_add["sigmaRp"], 
                    sigmaS0=prrs_add["sigmaS0"], sigmaSd=prrs_add["sigmaSd"], 
                    sigmaSp=prrs_add["sigmaSp"])

priorSn.add <- 
  list(list(d0=prrs_add["delB0S.1"], dp=prrs_add["delPhS.1"], dd=prrs_add["delDS.1"]),
       list(d0=prrs_add["delB0S.2"], dp=prrs_add["delPhS.2"], dd=prrs_add["delDS.2"]),
       list(d0=prrs_add["delB0S.3"], dp=prrs_add["delPhS.3"], dd=prrs_add["delDS.3"]),
       list(d0=prrs_add["delB0S.4"], dp=prrs_add["delPhS.4"], dd=prrs_add["delDS.4"]),
       list(d0=prrs_add["delB0S.5"], dp=prrs_add["delPhS.5"], dd=prrs_add["delDS.5"]),
       list(d0=prrs_add["delB0S.6"], dp=prrs_add["delPhS.6"], dd=prrs_add["delDS.6"]))

priorSn.add2020 <- priorSn.add ## save for model evaluation

### 2. nested
prrs_nstd <- rvsims(as.matrix(as.data.frame(rstan::extract(fit2keep_ann_2019N))))
## feeding to input function "priorYr_rv"
priorYr.nstd <- list(B0=prrs_nstd["B0"], De=prrs_nstd["De"], Ph=prrs_nstd["Ph"], 
                     sigmaR0=prrs_nstd["sigmaR0"], sigmaRd=prrs_nstd["sigmaRd"], 
                     sigmaRp=prrs_nstd["sigmaRp"])
priorSn.nstd <- list(sigmaS0=prrs_add["sigmaS0"], sigmaSd=prrs_add["sigmaSd"], 
                     sigmaSp=prrs_add["sigmaSp"])

### 3. Biweekly
prrs_2w <- rvsims(as.matrix(as.data.frame(rstan::extract(fit2keep_ann_2019W))))
## feeding to input function "priorYr_rv"
priorYr.2w <- list(B0=prrs_2w["B0"], De=prrs_2w["De"], Ph=prrs_2w["Ph"], 
                   sigmaR0=prrs_2w["sigmaR0"], sigmaRd=prrs_2w["sigmaRd"], 
                   sigmaRp=prrs_2w["sigmaRp"]) 
priorYr.2w <- list(sigmaS0=prrs_2w["sigmaS0"], sigmaSd=prrs_2w["sigmaSd"], 
                   sigmaSp=prrs_2w["sigmaSp"])
## sigmaRs needed for annual updating, but not for within a year

## 4. No seasons
prrs_nsn <- rvsims(as.matrix(as.data.frame(rstan::extract(fitNSN2019, permute=T))))
priors.nsn <- list(B0=prrs_nsn["mu0"], De=prrs_nsn["muD"], Ph=prrs_nsn["muP"], 
                   sigma0=prrs_nsn["sigma0"], sigmad=prrs_nsn["sigmad"], 
                   sigmap=prrs_nsn["sigmap"])

## processing data
load(paste(saveDIR, "eriedata2021.RData", sep="/"))
tmpdata <- eriedata2020
temp<-(!is.na(tmpdata$part_microcystin))&(!is.na(tmpdata$Chla))&(tmpdata$Chla>0)& 
  (tmpdata$part_microcystin>0)&(tmpdata$Month>=5 & tmpdata$Month<=10)
tmpdata <- tmpdata[temp,]
tmpdata$Mon_ord <- as.numeric(ordered(tmpdata$Month, levels=5:10))
tmpdata$Yr_ord <- as.numeric(ordered(tmpdata$Year))
tmpdata$sn_nstd <- as.numeric(ordered(paste(tmpdata$Year, tmpdata$Month, sep=":")))
## for biweek season
tmpdata$sn_nstd2w <- as.numeric(ordered(paste(tmpdata$Year, tmpdata$BiWeek, sep=":")))

seq_2020 <- seq_within_allY (data=tmpdata, add_yr0=priorYr.add, add_sn0=priorSn.add,
                             nstd_yr0=priorYr.nstd, nstd_sn0=priorSn.nstd, 
                             n2w_yr0=priorYr.2W, n2w_sn0 = priorSn.2W, 
                             nsn_yr0=priors.nsn)
save("seq_2020", file=paste(saveDIR, "seq2020.RData", sep="/"))
## load(paste(saveDIR, "seq2020.RData", sep="/"))
```


Year-end updating (2020)
```{R}
## updating priors for 2020:
### 1. Additive model
prrs_add <- rvsims(as.matrix(as.data.frame(rstan::extract(fit2keep_2019_add))))
## feeding to input function "priorYr_rv"
priorYr.add <- list(B0=prrs_add["B0"], De=prrs_add["De"], Ph=prrs_add["Ph"], 
                    sigmaR0=prrs_add["sigmaR0"], sigmaRd=prrs_add["sigmaRd"], 
                    sigmaRp=prrs_add["sigmaRp"], 
                    sigmaS0=prrs_add["sigmaS0"], sigmaSd=prrs_add["sigmaSd"], 
                    sigmaSp=prrs_add["sigmaSp"])

priorSn.add <- 
  list(list(d0=prrs_add["delB0S.1"], dp=prrs_add["delPhS.1"], dd=prrs_add["delDS.1"]),
       list(d0=prrs_add["delB0S.2"], dp=prrs_add["delPhS.2"], dd=prrs_add["delDS.2"]),
       list(d0=prrs_add["delB0S.3"], dp=prrs_add["delPhS.3"], dd=prrs_add["delDS.3"]),
       list(d0=prrs_add["delB0S.4"], dp=prrs_add["delPhS.4"], dd=prrs_add["delDS.4"]),
       list(d0=prrs_add["delB0S.5"], dp=prrs_add["delPhS.5"], dd=prrs_add["delDS.5"]),
       list(d0=prrs_add["delB0S.6"], dp=prrs_add["delPhS.6"], dd=prrs_add["delDS.6"]))


### 2. nested
prrs_nstd <- rvsims(as.matrix(as.data.frame(fit2keep_ann_2019N))) 
## feeding to input function "priorYr_rv"
priorYr.nstd <- list(B0=prrs_nstd["B0"], De=prrs_nstd["De"], Ph=prrs_nstd["Ph"], 
                     sigmaR0=prrs_nstd["sigmaR0"], sigmaRd=prrs_nstd["sigmaRd"], 
                     sigmaRp=prrs_nstd["sigmaRp"])
priorSn.nstd <- list(sigmaS0=prrs_add["sigmaS0"], sigmaSd=prrs_add["sigmaSd"], 
                     sigmaSp=prrs_add["sigmaSp"])

### 3. Biweekly
prrs_2w <- rvsims(as.matrix(as.data.frame(fit2keep_ann_2019W))) 
## feeding to input function "priorYr_rv"
priorYr.2w <- list(B0=prrs_2w["B0"], De=prrs_2w["De"], Ph=prrs_2w["Ph"], 
                   sigmaR0=prrs_2w["sigmaR0"], sigmaRd=prrs_2w["sigmaRd"], 
                   sigmaRp=prrs_2w["sigmaRp"]) 
priorYr.2w <- list(sigmaS0=prrs_2w["sigmaS0"], sigmaSd=prrs_2w["sigmaSd"], 
                   sigmaSp=prrs_2w["sigmaSp"])
## sigmaRs needed for annual updating, but not for within a year

### 4. No season using `stan_mod_cens` (line 97)
prrs_nsn <- rvsims(as.matrix(as.data.frame(rstan::extract(fitNSN2019, permute=T))))
priors.nsn <- list(B0=prrs_nsn["mu0"], De=prrs_nsn["muD"], Ph=prrs_nsn["muP"], 
                   sigma0=prrs_nsn["sigma0"], sigmad=prrs_nsn["sigmad"], 
                   sigmap=prrs_nsn["sigmap"])

## 2020 year end updating

tmpdata <- eriedata2020
temp<-(!is.na(tmpdata$part_microcystin))&(!is.na(tmpdata$Chla))&(tmpdata$Chla>0)& 
  (tmpdata$part_microcystin>0)&(tmpdata$Month>=5 & tmpdata$Month<=10)
tmpdata <- tmpdata[temp,]
tmpdata$Mon_ord <- as.numeric(ordered(tmpdata$Month, levels=5:10))
tmpdata$Yr_ord <- as.numeric(ordered(tmpdata$Year))
tmpdata$sn_nstd <- as.numeric(ordered(paste(tmpdata$Year, tmpdata$Month, sep=":")))
## for biweek season
tmpdata$sn_nstd2w <- as.numeric(ordered(paste(tmpdata$Year, tmpdata$BiWeek, sep=":")))

## no seasons
input.to.stan <- stan.in_seq_nsn(infile=tmpdata, priorYr_rv = priors.nsn)
fitNSN2020 <- sampling(stan_mod_cens,
            			     data = input.to.stan$data,
	                     init = input.to.stan$inits,
                       pars = input.to.stan$para,
                       iter = niters, thin=nthin, 
                       chains=input.to.stan$n.chains) #, 
##                   control = list(adapt_delta = 0.9, max_treedepth=15))
## additive
input.to.stan <- stan.in_seqA_add(infile=tmpdata, x="Chla", y="part_microcystin",
                              n.chains=nchains, priorYr_rv=priorYr.add, 
                              priorSn_rv=priorSn.add)

fit2keep_2020_add <- sampling(stan_mod_ann_add, 
                              data = input.to.stan$data,
                              init=input.to.stan$inits,
                              pars = input.to.stan$para,
                              iter=niters, thin=nthin,
                              chains=input.to.stan$n.chains)

print(fit2keep_2020_add)
plot(fit2keep_2020_add)

## nested (month)
input.to.stan <- stan.in_seqA_nstd(infile=tmpdata, x="Chla", y="part_microcystin",
                              n.chains=nchains, priorYr_rv=priorYr.nstd, priorSn_rv=priorSn.nstd,biweek=F)

fit2keep_ann_2020N <- sampling(stan_mod_ann_nstd, 
                               data = input.to.stan$data,
                               init=input.to.stan$inits,  
                               pars = input.to.stan$para,
                               iter=niters, thin=nthin,
                               chains=input.to.stan$n.chains)#)

print(fit2keep_ann_2020N)
plot(fit2keep_ann_2020N)
## 2w
input.to.stan <- stan.in_seqA_nstd(infile=tmpdata, x="Chla", y="part_microcystin",
                              n.chains=nchains, priorYr_rv=priorYr.nstd, 
                              priorSn_rv=priorSn.nstd,biweek=T)

fit2keep_ann_2020W <- sampling(stan_mod_ann_nstd,  data = input.to.stan$data,
                               init=input.to.stan$inits,
                               pars = input.to.stan$para,
                               iter=niters, thin=nthin,
                               chains=input.to.stan$n.chains)#)

print(fit2keep_ann_2020W)
plot(fit2keep_ann_2020W)

save("fitNSN2020", "fit2keep_2020_add", "fit2keep_ann_2020N", "fit2keep_ann_2020W", 
     file=paste(saveDIR, "annual2020.RData", sep="/"))
## load(paste(saveDIR, "annual2020.RData", sep="/"))
```

Now running sequential updating within a year (2021):

```{R}

## updating priors for 2021: Extracting priors

### 1. Additive model
prrs_add <- rvsims(as.matrix(as.data.frame(rstan::extract(fit2keep_2020_add))))
## feeding to input function "priorYr_rv"
priorYr.add <- list(B0=prrs_add["B0"], De=prrs_add["De"], Ph=prrs_add["Ph"], 
                   sigmaR0=prrs_add["sigmaR0"], sigmaRd=prrs_add["sigmaRd"], 
                   sigmaRp=prrs_add["sigmaRp"], 
                   sigmaS0=prrs_add["sigmaS0"], sigmaSd=prrs_add["sigmaRd"], 
                   sigmaSp=prrs_add["sigmaSp"])
priorSn.add <- list(list(d0=prrs_add["delB0S.1"], dp=prrs_add["delPhS.1"], dd=prrs_add["delDS.1"]),
                    list(d0=prrs_add["delB0S.2"], dp=prrs_add["delPhS.2"], dd=prrs_add["delDS.2"]),
                    list(d0=prrs_add["delB0S.3"], dp=prrs_add["delPhS.3"], dd=prrs_add["delDS.3"]),
                    list(d0=prrs_add["delB0S.4"], dp=prrs_add["delPhS.4"], dd=prrs_add["delDS.4"]),
                    list(d0=prrs_add["delB0S.5"], dp=prrs_add["delPhS.5"], dd=prrs_add["delDS.5"]),
                    list(d0=prrs_add["delB0S.6"], dp=prrs_add["delPhS.6"], dd=prrs_add["delDS.6"]))

priorSn.add2021 <- priorSn.add ## save for model evaluation

### 2. nested
prrs_nstd <- rvsims(as.matrix(as.data.frame(rstan::extract(fit2keep_ann_2020N)))) 
## feeding to input function "priorYr_rv"
priorYr.nstd <- list(B0=prrs_nstd["B0"], De=prrs_nstd["De"], Ph=prrs_nstd["Ph"], 
                     sigmaR0=prrs_nstd["sigmaR0"], sigmaRd=prrs_nstd["sigmaRd"], 
                     sigmaRp=prrs_nstd["sigmaRp"])
priorSn.nstd <- list(sigmaS0=prrs_add["sigmaS0"], sigmaSd=prrs_add["sigmaSd"], 
                     sigmaSp=prrs_add["sigmaSp"])

### 3. Biweekly
prrs_2w <- rvsims(as.matrix(as.data.frame(rstan::extract(fit2keep_ann_2020W))))
## feeding to input function "priorYr_rv"
priorYr.2w <- list(B0=prrs_2w["B0"], De=prrs_2w["De"], Ph=prrs_2w["Ph"], 
                   sigmaR0=prrs_2w["sigmaR0"], sigmaRd=prrs_2w["sigmaRd"], 
                   sigmaRp=prrs_2w["sigmaRp"]) 
priorYr.2w <- list(sigmaS0=prrs_2w["sigmaS0"], sigmaSd=prrs_2w["sigmaSd"], 
                   sigmaSp=prrs_2w["sigmaSp"])

### 4. No season using `stan_mod_cens` (line 97)
prrs_nsn <- rvsims(as.matrix(as.data.frame(rstan::extract(fitNSN2020))))
priors.nsn <- list(B0=prrs_nsn["mu0"], De=prrs_nsn["muD"], Ph=prrs_nsn["muP"], 
                   sigma0=prrs_nsn["sigma0"], sigmad=prrs_nsn["sigmad"], 
                   sigmap=prrs_nsn["sigmap"])
## 2021
## processing data
tmpdata <- eriedata2021
temp<-(!is.na(tmpdata$part_microcystin))&(!is.na(tmpdata$Chla))&(tmpdata$Chla>0)& 
  (tmpdata$part_microcystin>0)&(tmpdata$Month>=5 & tmpdata$Month<=10)
tmpdata <- tmpdata[temp,]
tmpdata$Mon_ord <- as.numeric(ordered(tmpdata$Month, levels=5:10))
tmpdata$Yr_ord <- as.numeric(ordered(tmpdata$Year))
tmpdata$sn_nstd <- as.numeric(ordered(paste(tmpdata$Year, tmpdata$Month, sep=":")))
## for biweek season
tmpdata$sn_nstd2w <- as.numeric(ordered(paste(tmpdata$Year, tmpdata$BiWeek, sep=":")))

seq_2021 <- seq_within_allY (data=tmpdata, add_yr0=priorYr.add, add_sn0=priorSn.add,
                             nstd_yr0=priorYr.nstd, nstd_sn0=priorSn.nstd, 
                             n2w_yr0=priorYr.2W, n2w_sn0 = priorSn.2W, 
                             nsn_yr0=priors.nsn)
save("seq_2021", file=paste(saveDIR, "seq2021.RData", sep="/"))
## load(paste(saveDIR, "seq2021.RData", sep="/"))
```

Year-end updating (2021)
```{R}
## updating priors for 2021:
### 1. Additive model
### 1. Additive model
prrs_add <- rvsims(as.matrix(as.data.frame(rstan::extract(fit2keep_2020_add))))
## feeding to input function "priorYr_rv"
priorYr.add <- list(B0=prrs_add["B0"], De=prrs_add["De"], Ph=prrs_add["Ph"], 
                   sigmaR0=prrs_add["sigmaR0"], sigmaRd=prrs_add["sigmaRd"], 
                   sigmaRp=prrs_add["sigmaRp"], 
                   sigmaS0=prrs_add["sigmaS0"], sigmaSd=prrs_add["sigmaRd"], 
                   sigmaSp=prrs_add["sigmaSp"])
priorSn.add <- list(list(d0=prrs_add["delB0S.1"], dp=prrs_add["delPhS.1"], dd=prrs_add["delDS.1"]),
                    list(d0=prrs_add["delB0S.2"], dp=prrs_add["delPhS.2"], dd=prrs_add["delDS.2"]),
                    list(d0=prrs_add["delB0S.3"], dp=prrs_add["delPhS.3"], dd=prrs_add["delDS.3"]),
                    list(d0=prrs_add["delB0S.4"], dp=prrs_add["delPhS.4"], dd=prrs_add["delDS.4"]),
                    list(d0=prrs_add["delB0S.5"], dp=prrs_add["delPhS.5"], dd=prrs_add["delDS.5"]),
                    list(d0=prrs_add["delB0S.6"], dp=prrs_add["delPhS.6"], dd=prrs_add["delDS.6"]))

### 2. nested
prrs_nstd <- rvsims(as.matrix(as.data.frame(rstan::extract(fit2keep_ann_2020N))))
## feeding to input function "priorYr_rv"
priorYr.nstd <- list(B0=prrs_nstd["B0"], De=prrs_nstd["De"], Ph=prrs_nstd["Ph"], 
                     sigmaR0=prrs_nstd["sigmaR0"], sigmaRd=prrs_nstd["sigmaRd"], 
                     sigmaRp=prrs_nstd["sigmaRp"])
priorSn.nstd <- list(sigmaS0=prrs_add["sigmaS0"], sigmaSd=prrs_add["sigmaSd"], 
                     sigmaSp=prrs_add["sigmaSp"])

### 3. Biweekly
prrs_2w <- rvsims(as.matrix(as.data.frame(rstan::extract(fit2keep_ann_2020W))))
## feeding to input function "priorYr_rv"
priorYr.2w <- list(B0=prrs_2w["B0"], De=prrs_2w["De"], Ph=prrs_2w["Ph"], 
                   sigmaR0=prrs_2w["sigmaR0"], sigmaRd=prrs_2w["sigmaRd"], 
                   sigmaRp=prrs_2w["sigmaRp"]) 
priorYr.2w <- list(sigmaS0=prrs_2w["sigmaS0"], sigmaSd=prrs_2w["sigmaSd"], 
                   sigmaSp=prrs_2w["sigmaSp"])

### 4. No season using `stan_mod_cens` (line 97)
prrs_nsn <- rvsims(as.matrix(as.data.frame(rstan::extract(fitNSN2020))))
priors.nsn <- list(B0=prrs_nsn["mu0"], De=prrs_nsn["muD"], Ph=prrs_nsn["muP"], 
                   sigma0=prrs_nsn["sigma0"], sigmad=prrs_nsn["sigmad"], 
                   sigmap=prrs_nsn["sigmap"])
## 2021 year end updating

tmpdata <- eriedata2021
temp<-(!is.na(tmpdata$part_microcystin))&(!is.na(tmpdata$Chla))&(tmpdata$Chla>0)& 
  (tmpdata$part_microcystin>0)&(tmpdata$Month>=5 & tmpdata$Month<=10)
tmpdata <- tmpdata[temp,]
tmpdata$Mon_ord <- as.numeric(ordered(tmpdata$Month, levels=5:10))
tmpdata$Yr_ord <- as.numeric(ordered(tmpdata$Year))
tmpdata$sn_nstd <- as.numeric(ordered(paste(tmpdata$Year, tmpdata$Month, sep=":")))
tmpdata$sn_nstd2w <- as.numeric(ordered(paste(tmpdata$Year, tmpdata$BiWeek, sep=":")))

## no seasons
input.to.stan <- stan.in_seq_nsn(infile=tmpdata, priorYr_rv = priors.nsn)
fitNSN2021 <- sampling(stan_mod_cens,
            			     data = input.to.stan$data,
	                     init = input.to.stan$inits,
                       pars = input.to.stan$para,
                       iter = niters, thin=nthin, 
                       chains=input.to.stan$n.chains) #, 
##                   control = list(adapt_delta = 0.9, max_treedepth=15))
## additive
input.to.stan <- stan.in_seqA_add(infile=tmpdata, x="Chla", y="part_microcystin",
                              n.chains=nchains, priorYr_rv=priorYr.add, 
                              priorSn_rv=priorSn.add)

fit2keep_2021_add <- sampling(stan_mod_ann_add, 
                              data = input.to.stan$data,
                              init=input.to.stan$inits,
                              pars = input.to.stan$para,
                              iter=niters, thin=nthin,
                              chains=input.to.stan$n.chains)

print(fit2keep_2021_add)
plot(fit2keep_2021_add)

## nested (month)
input.to.stan <- stan.in_seqA_nstd(infile=tmpdata, x="Chla", y="part_microcystin",
                                   n.chains=nchains, priorYr_rv=priorYr.nstd, 
                                   priorSn_rv=priorSn.nstd,biweek=F)

fit2keep_ann_2021N <- sampling(stan_mod_ann_nstd, 
                               data = input.to.stan$data,
                               init=input.to.stan$inits,  
                               pars = input.to.stan$para,
                               iter=niters, thin=nthin,
                               chains=input.to.stan$n.chains)#)

print(fit2keep_ann_2021N)
plot(fit2keep_ann_2021N)
## 2w
input.to.stan <- stan.in_seqA_nstd(infile=tmpdata, x="Chla", y="part_microcystin",
                              n.chains=nchains, priorYr_rv=priorYr.nstd, priorSn_rv=priorSn.nstd,biweek=T)

fit2keep_ann_2021W <- sampling(stan_mod_ann_nstd,  data = input.to.stan$data,
                               init=input.to.stan$inits,
                               pars = input.to.stan$para,
                               iter=niters, thin=nthin,
                               chains=input.to.stan$n.chains)#)

print(fit2keep_ann_2021W)
plot(fit2keep_ann_2021W)

save("fitNSN2021", "fit2keep_2021_add", "fit2keep_ann_2021N", "fit2keep_ann_2021W", 
     file=paste(saveDIR, "annual2021.RData", sep="/"))
save("priorSn.add2019", "priorSn.add2020", "priorSn.add2021", file=paste(saveDIR, "prior19_21.RData", sep="/"))

```
